---
title: "wearable devices"
author: "me"
date: "2/22/2021"
output:
  html_document:
    df_print: paged
---


setup environment and load data

```{r}
library(caret)
library(e1071)
library(randomForest)
library(psych)
library(gbm)

train_data <- read.csv('pml-training.csv')
test_data <- read.csv('pml-testing.csv')
```

preprocessing

- remove columns that do not pertain to exercise

```{r}
scoped_train <- train_data[,-1:-7]
scoped_test <- test_data[,-1:-7]
```

split data for train and test

```{r}
set.seed(1)
train_idx <- createDataPartition(scoped_train$classe, p = 0.7, list = FALSE)
validate_df <- scoped_train[-train_idx, ]
training_df <- scoped_train[train_idx, ]
```

remove near zero var

```{r}
nzv <- nearZeroVar(training_df, saveMetrics = T)
features_to_keep <- row.names(nzv[nzv$nzv == FALSE, ])
training_df <- training_df[, features_to_keep]
```

remove all na

```{r}
training_df <- training_df[, colSums(is.na(training_df)) == 0]
```

apply pca to reduce dimension

```{r}
fa.parallel(training_df[,1:52], main="Scree plot with parallel analysis")
```

remove factors with eigenvalues < 1

```{r}
pca_scope <- preProcess(training_df[,1:52], method = "pca", pcaComp = 13)
train_to_model <- predict(pca_scope, training_df)
test_to_model <- predict(pca_scope, validate_df)
```

fit random forest model

```{r}
start_time <- Sys.time()
set.seed(123)
rf_model <- train(classe ~. , data = train_to_model, method = 'rf', nodesize = 7, ntree = 30)
end_time <- Sys.time()
end_time - start_time
```

predict with validation set

```{r}
rf_preds <- predict(rf_model, newdata = test_to_model)
confusionMatrix(rf_preds,as.factor(test_to_model$classe))$overall[1]
```

fit gradient boosted model

```{r}
start_time <- Sys.time()
gbm_model <- train(classe ~., data = train_to_model, method = 'gbm', tuneGrid = expand.grid(
  n.trees = 100, 
  interaction.depth = 1, 
  shrinkage = .1,
  n.minobsinnode = 10),
  verbose = F)
end_time <- Sys.time()
end_time - start_time
```

predict with validation set

```{r}
gbm_preds <- predict(gbm_model, newdata = test_to_model)
confusionMatrix(gbm_preds,as.factor(test_to_model$classe))$overall[1]
```

predict test using best model (random forest)

```{r}
test_final <- predict(pca_scope, scoped_test)
predict(rf_model, newdata = test_final)
```


